Make any stupid plot you want

### 2021-01-27 draft/sections/math.tex
rewrote intro again

\section{Topological Artist Model}
\note{can change this all to 3rd person passive later}\\
In this section we introduce a mathematical description of the visualization pipeline where artist $\mathscr{A}$ functions transform data space $\mathscr{E}$ to an intermediate representation in a prerendered graphic space $\mathscr{H}$.

@@

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
\note{yes, I know i may need to change this U to like $\mathcal{U}$, spivak uses U}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of types $\{T_0, \ldots, T_n\} \in DT$ where $DT$ are the data types of the variables in the dataset. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U{T_i} \ldots\times U_{T_n}
\end{equation}

@@

For example, the data in figure~\ref{fig:fiber_example_plane} is a pair of times and \textdegree C temperature measurements taken at those times. Time is a positive number ot type \texttt{detetime} which can be resolved to positive floats $U_{\texttt{datetime}}= \mathbb{R}^+$. Temperature values are real numbers $U_{\texttt{float}} = \mathbb{R}$. The fiber is 
\begin{equation}
    F =  \mathbb{R}^{+} \times \mathbb{R} 
\end{equation} 
where $V_0 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$ and $V_1 = (temperature,\, \texttt{float},\, \mathbb{R})$. In figure~\ref{fig:fiber_example_cube}, temperature is replaced with wind. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(s,d) \in \mathbb{R}^{2} \mid  0\leq s,\, 0 \leq d \leq 360\}$ . Therefore, the fiber is 
\begin{equation}
    F = \mathbb{R}^{+} \times \mathbb{R}^2
\end{equation} 
such that $V_1 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

@@

\subsubsection{Abstract Display $D$}
\label{sec:graphic_fiber}
The fiber $D$ is specified in terms of the target display space, for example a 2D screen or 3D printer. In this work, we assume a 2D opaque image $D=\mathbb{R}^5$ with elements in $D$ of the form
\begin{equation}
    d = (x, y, r, g, b)
\end{equation}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $D=R^{7}$ such that $d=(x, y, z, r, g, b, a)$. The location coordinates $x$ and $y$ are defined in terms of the display, while the $(r,g,b)$ values are filled in via a lookup on $S$.

@@

\subsubsection{Renderable Glyph $\rho$}
\label{sec:graphic_section}

A section $\rho: S\rightarrow H$ defines a piece of the graphical representation of the data. The physical display coordinates in $D$ are not equivalent to the mathematical coordinates in $S$ such that many points $s \in S$ might map to a single $d\in D$ and many $d$ might go into a visible section of a displayed graphic. Therefore, we define the bounding box $b=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ in the display space $D$ that is a piece of the rendered graphic.

@@

\subsection{Artist}
\label{sec:artist}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/dar.png}
    \caption{The artist $A$ converts an section of data $\tau$ into a section of visual equivalents $\mu$ that are then composed into a piece of a glyph defined by the graphic section $\rho$
    \note{might need to label this as E', V, H}}
    \label{fig:artist_stages}
\end{figure}

@@

\subsubsection {Visual Fiber Bundle $V$}
The visual fiber bundle ($V$, $K$, $\pi$, $P$) has section $\mu: V \rightarrow K$ that resolves to a visual variable \cite{bertinIIPropertiesGraphic2011,munznerMarksChannels,} in fiber $P$. The fiber space $P$ is defined in terms of the parameters of the visualization specification and the values those parameters resolve to - for example APT\cite{mackinlayAutomatingDesignGraphical1986} or Vega\cite{satyanarayanDeclarativeInteractionDesign2014}.

\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\nu_{i}}$                      & $\bm{\mu_{i}}$                                                            & $\bm{codomain(\nu_{i})}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $\{f_{0}, \ldots, f_{n}\} \times (\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber $P$ for a visualization implemented in Matplotlib}
    \label{tab:mpl_visual_variable_fiber}
\end{table}

@@

\subsubsection{Assembling Marks}
\note{HAVE NOT REACHED HERE TO EDIT YET}
\label{sec:artist_q}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}
    \label{fig:artist_q}
    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$.}
    %%maybe change to \tau, \mu, \rho  
\end{figure}

>>>>

\section{Topological Artist Model}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/dar.png}
    \caption{Visualizaiton is the process of converting data into visual characteristics that are then assembled into a graphic}
    \label{fig:artist_stages}
\end{figure}

\note{can change this all to 3rd person passive later}\\
%% add sentence or two about visualization as structure preserving map
In this section we introduce a mathematical description of the visualization pipeline where artist $\mathscr{A}$ functions transform data space $\mathscr{E}$ to an intermediate representation in a prerendered graphic space $\mathscr{H}$.

@@

\subsubsection{Variables: Fiber Space $F$}
\label{sec:data_fiber}
\note{yes, I know i may need to change this U to like $\mathcal{U}$, spivak uses U}
The fiber is a topological space that is the set of possible values of the data; the values themselves can be any dimension and type and have any continuity. We use Spivak's description of simplicial database schemas \cite{spivakSIMPLICIALDATABASES} as the basis of our fiber space because he binds the components of the fiber to variable names and types. Spivak constructs a set $U$ that is the disjoint union of all possible objects of types $\{T_0, \ldots, T_n\} \in DT$ where $DT$ are the data types of the variables in the dataset. He then defines the single variable set $U_\sigma$ 
\begin{equation}
    \label{eq:data_types}
\begin{tikzcd}
    U_{\sigma} \arrow[r] \arrow[d, "\pi_{\sigma}"'] & U \arrow[d, "\pi"] \\
    C \arrow[r, "\sigma"']                          & \textbf{DT}       
\end{tikzcd}
\end{equation}
which is $U$ restricted to objects of type $T$ bound to variable name $c$. Given $U_{\sigma}$, the fiber for a one variable dataset is
\begin{equation}
    F = U_{\sigma(c)} = U_{T} 
\end{equation}
where $\sigma$ is the schema binding variable name $c$ to its datatype $T$. A dataset with multiple variables has a fiber that is the cartesian cross product of $U_{\sigma}$ applied to all the columns:
\begin{equation}
F = U_{T_1}\times \ldots U_{T_i} \ldots\times U_{T_n}
\end{equation}

@@

For example, the data in figure~\ref{fig:fiber_example_plane} is a pair of times and \textdegree C temperature measurements taken at those times. Time is a positive number of type \texttt{detetime} which can be resolved to positive floats $U_{\texttt{datetime}}= \mathbb{R}^+$. Temperature values are real numbers $U_{\texttt{float}} = \mathbb{R}$. The fiber is 
\begin{equation}
    F =  \mathbb{R}^{+} \times \mathbb{R} 
\end{equation} 
where $V_0 = (time,\, \texttt{datetime},\, \mathbb{R}^+)$ and $V_1 = (temperature,\, \texttt{float},\, \mathbb{R})$. In figure~\ref{fig:fiber_example_cube}, temperature is replaced with wind. This wind variable is of type \texttt{wind} and has two components speed and direction $\{(s,d) \in \mathbb{R}^{2} \mid  0\leq s,\, 0 \leq d \leq 360\}$ . Therefore, the fiber is 
\begin{equation}
    F = \mathbb{R}^{+} \times \mathbb{R}^2
\end{equation} 
such that $V_1 = (wind,\, \texttt{wind},\, \mathbb{R}^{2})$

@@

\subsubsection{Idealized Display $D$}
\label{sec:graphic_fiber}
% infinite resolution version of D
The fiber $D$ is specified in terms of the target display space, for example a 2D screen or 3D printer. In this work, we assume a 2D opaque image $D=\mathbb{R}^5$ with elements in $D$ of the form
\begin{equation}
    d = (x, y, r, g, b)
\end{equation}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $D=R^{7}$ such that $d=(x, y, z, r, g, b, a)$. The location coordinates $x$ and $y$ are defined in terms of the display, while the $(r,g,b)$ values are filled in via a lookup on $S$.

@@

\subsubsection{Renderable Glyph $\rho$}
\label{sec:graphic_section}
%pick a pixel which is the bounding box
A section $\rho: S\rightarrow H$ defines a piece of the graphical representation of the data. The physical display coordinates in $D$ are not equivalent to the mathematical coordinates in $S$ such that many points $s \in S$ might map to a single $d\in D$ and many $d$ might go into a visible section of a displayed graphic. Therefore, we define the bounding box $b=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ in the display space $D$ that is a piece of the rendered graphic.

@@

\subsection{Artist}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/path_of_q}

    \caption{The $\nu$ functions convert data $E$ to visual $V$. $Q$ assembles the different types of visual parameters $V_{i}$ into a graphic in $H$. $Q\circ\mu(\xi^{-1}J)$ forms a visual mark by applying $Q$ to a region mapped to connected components $J \subset K$. \note{maybe change to $\tau, \mu, \rho$ }} 
    \label{fig:artist_q}
\end{figure}

@@

\subsubsection {Visual Fiber Bundle $V$}
The visual fiber bundle ($V$, $K$, $\pi$, $P$) has section $\mu: V \rightarrow K$ that resolves to a visual variable \cite{bertinIIPropertiesGraphic2011,munznerMarksChannels,} in fiber $P$. The fiber space $P$ is defined in terms of the parameters of the visualization specification and the values those parameters resolve to - for example ggplot \cite{wickhamGgplot2ElegantGraphics2016a}, D3\cite{bostockDataDrivenDocuments2011
\begin{table}[H]
    \renewcommand{\arraystretch}{2}
    \begin{tabulary}{\textwidth}{|l|L|l|}\hline
     $\bm{\nu_{i}}$                      & $\bm{\mu_{i}}$                                                            & $\bm{codomain(\nu_{i})}$  \\ \hline                                              
    position                    & x, y, z, theta, r                                                          & $\mathbb{R}$   \\ \hline
    size                        & linewidth, markersize                                            & $\mathbb{R}^{+}$   \\ \hline
    shape                       & markerstyle                                                      & $\{f_{0}, \ldots, f_{n}\}$ \\ \hline
    color                       & color, facecolor, markerfacecolor, edgecolor  & $\mathbb{R}^{4}$ \\ \hline
    \multirow{2}{*}{texture}    & hatch                                                            & $\mathbb{N}^{10}$\\\cline{2-3}
                                & linestyle                                                        & $\{f_{0}, \ldots, f_{n}\} \times (\mathbb{R}, \mathbb{R^+}^{n, n\%2=0})$ \\ \hline              
    \end{tabulary}
    \caption{Some possible components of the fiber $P$ for a visualization implemented in Matplotlib}
    \label{tab:mpl_visual_variable_fiber}
\end{table}

@@

\subsubsection{Assembling Marks}
\label{sec:artist_q}
